---
title: "The socioeconomic consequences of loneliness: Evidence from a nationally representative longitudinal study of young adults. Code review by K N Thompson"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
    number_sections: no
    highlight: monochrome
    theme: flatly
    code_folding: hide
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      results = 'asis')

options(bitmapType = 'quartz') # to render fonts better
```

```{r Clear global environment, include=FALSE}
remove(list = ls())
```

```{r Load packages, include=FALSE}
library(knitr)
library(haven)
library(lavaan)
library(ChiBarSq.DiffTest)
library(janitor)
library(tidyr)
library(tidyverse)
library(dplyr) #conflicts with tidyverse for e.g. rename and row_number
```

```{r source the data file path, include=FALSE}
#source raw data directory: data_raw and data included
source("file_path.R")
```

# Functions

```{r functions}
# Table of model fit 
table.model.fit <- function(model){
  model.fit <- as.data.frame(t(as.data.frame(model$FIT))) %>%
    dplyr::select(chisq, df, chisq.scaled, cfi.robust, tli.robust, aic, bic, bic2, rmsea.robust, rmsea.ci.lower.robust, rmsea.ci.upper.robust, srmr) #can only be used with "MLR" estimator
  return(model.fit)
}

# Table of regression and correlation (standardised covariance) coefficients
table.model.coef <- function(model, type, constraints){
  if (type == "RICLPM" & constraints == "No"){
    model.coef <- as.tibble(model$PE[c(13:32),]) %>% 
      dplyr::select(lhs, op, rhs, std.all, pvalue) %>%
      mutate_if(is.numeric, round, 3)
    return(model.coef)
    
  } else if(type == "RICLPM" & constraints == "Yes"){
    model.coef <- as.tibble(model$PE[c(13:32),]) %>% 
      dplyr::select(lhs, op, rhs, label, std.all, pvalue) %>%
      mutate_if(is.numeric, round, 3)
    return(model.coef)
    
  } else if(type == "CLPM" & constraints == "No"){
    model.coef <- as.tibble(model$PE[c(1:17),]) %>% 
      dplyr::select(lhs, op, rhs, std.all, pvalue) %>%
      mutate_if(is.numeric, round, 3)
    return(model.coef)
    
  } else if(type == "CLPM" & constraints == "Yes"){
    model.coef <- as.tibble(model$PE[c(1:17),]) %>% 
      dplyr::select(lhs, op, rhs, label, std.all, pvalue) %>%
      mutate_if(is.numeric, round, 3)
    return(model.coef)
    
  } else {model.coef <- NULL}
}
```

# Data prep

Loneliness variables:

* lonelye12
* lonelye18
* lonelysm2e

Subjective social status:

* sta04ec12
* sicountrye18
* psstdysm2er

```{r read in dta data file}
dat.raw <- read_dta(paste0(data.raw_path))

dat <- dat.raw %>%
  dplyr::select(atwinid,
         sampsex,
         seswq35,
         sethnic,
         lonelye12,
         lonelye18,
         lonelysm2e,
         sta04ec12,
         sicountrye18,
         psstdysm2er
  )

colnames(dat)
```

# CLPM

Stata code (Standardised co-efficients, maximum likelihood with missing):

sem (lonelye12 -> lonelye18, ) # done
    (lonelye12 -> sicountrye18, ) # done
    (sta04ec12 -> lonelye18, ) # done
    (sta04ec12 -> sicountrye18, ) # done
    (lonelye18 -> lonelysm2e, ) # done
    (lonelye18 -> psstdysm2er, ) # done
    (sicountrye18 -> lonelysm2e, ) # done
    (sicountrye18 -> psstdysm2er, ), # done
      method(mlmv) 
      vce(cluster familyid) 
      standardized cov( lonelye12*sta04ec12 e.lonelye18*e.sicountrye18 e.lonelysm2e*e.psstdysm2er) 
      nocapslatent
// N = 2,195

```{r specify CLPM}
CLPM <- '
  # Estimate the lagged effects between the observed variables
  lonelye18 + sicountrye18 ~ lonelye12 + sta04ec12 # loneliness and subsocstatus at 12 predicting them at 18
  lonelysm2e + psstdysm2er ~ lonelye18 + sicountrye18 # loneliness and subsocstatus at 18 predicting them at 26

  # Estimate the covariance between the observed variables at the first wave 
  lonelye12 ~~ sta04ec12 # Covariance
  
  # Estimate the covariances between the residuals of the observed variables
  lonelye18 ~~ sicountrye18
  lonelysm2e ~~ psstdysm2er
  
  # Estimate the (residual) variance of the observed variables.
  lonelye12 ~~ lonelye12 # Variances
  sta04ec12 ~~ sta04ec12 
  lonelye18 ~~ lonelye18 # Residual variances
  sicountrye18 ~~ sicountrye18 
  lonelysm2e ~~ lonelysm2e 
  psstdysm2er ~~ psstdysm2er 
'
```

```{r fit CLPM}
CLPM.fit <- lavaan(CLPM, 
                   data = dat, 
                   missing = 'ML',
                   meanstructure = TRUE, 
                   int.ov.free = TRUE,
                   se = "robust",
                   estimator = "MLR" #maximum likelihood with robust (Huber-White) standard errors and a scaled (Yuan-Bentler) and robust test statistic 
                   ) 

CLPM.fit.summary <- summary(CLPM.fit, 
                            fit.measures = TRUE,
                            standardized = TRUE)

#Table of model fit 
CLPM.fit.summary.fit <- table.model.fit(CLPM.fit.summary)
CLPM.fit.summary.fit

#Table of regression coefficients and covariances (concurrent associations)
CLPM.fit.summary.reg <- table.model.coef(model = CLPM.fit.summary, type = "CLPM", constraints = "No")
CLPM.fit.summary.reg
```

 Used 2195 observations, same as Stata. 
 
 All regression paths replicate those found in Stata. 
 
# Constrained CLPM - autoregressive 
 
```{r specify CLPM2}
CLPM2 <- '
  # Estimate the lagged effects between the observed variables
  lonelye18  ~ a*lonelye12 + sta04ec12 
  sicountrye18 ~ lonelye12 + b*sta04ec12 
  
  lonelysm2e ~ a*lonelye18 + sicountrye18 
  psstdysm2er ~ lonelye18 + b*sicountrye18 

  # Estimate the covariance between the observed variables at the first wave 
  lonelye12 ~~ sta04ec12 # Covariance
  
  # Estimate the covariances between the residuals of the observed variables
  lonelye18 ~~ sicountrye18
  lonelysm2e ~~ psstdysm2er
  
  # Estimate the (residual) variance of the observed variables.
  lonelye12 ~~ lonelye12 # Variances
  sta04ec12 ~~ sta04ec12 
  lonelye18 ~~ lonelye18 # Residual variances
  sicountrye18 ~~ sicountrye18 
  lonelysm2e ~~ lonelysm2e 
  psstdysm2er ~~ psstdysm2er 
'
```
 
```{r fit CLPM2}
CLPM2.fit <- lavaan(CLPM2, 
                   data = dat, 
                   missing = 'ML',
                   meanstructure = TRUE, 
                   int.ov.free = TRUE,
                   se = "robust",
                   estimator = "MLR" #maximum likelihood with robust (Huber-White) standard errors and a scaled (Yuan-Bentler) and robust test statistic 
                   ) 

CLPM2.fit.summary <- summary(CLPM2.fit, 
                            fit.measures = TRUE,
                            standardized = TRUE)

#Table of model fit 
CLPM2.fit.summary.fit <- table.model.fit(CLPM2.fit.summary)
CLPM2.fit.summary.fit

#Table of regression coefficients and covariances (concurrent associations)
CLPM2.fit.summary.reg <- table.model.coef(model = CLPM2.fit.summary, type = "CLPM", constraints = "Yes")
CLPM2.fit.summary.reg
```

```{r LRT for CLPMcomb and CLPMcomb2}
lavTestLRT(CLPM.fit, CLPM2.fit, method = "satorra.bentler.2010")
```
Not significantly worse fit - can constrain autoregressive paths. 

# Constrained CLPM - autoregressive + crosslag
 
```{r specify CLPM3}
CLPM3 <- '
  # Estimate the lagged effects between the observed variables
  lonelye18  ~ a*lonelye12 + c*sta04ec12 
  sicountrye18 ~ d*lonelye12 + b*sta04ec12 
  
  lonelysm2e ~ a*lonelye18 + c*sicountrye18 
  psstdysm2er ~ d*lonelye18 + b*sicountrye18 

  # Estimate the covariance between the observed variables at the first wave 
  lonelye12 ~~ sta04ec12 # Covariance
  
  # Estimate the covariances between the residuals of the observed variables
  lonelye18 ~~ sicountrye18
  lonelysm2e ~~ psstdysm2er
  
  # Estimate the (residual) variance of the observed variables.
  lonelye12 ~~ lonelye12 # Variances
  sta04ec12 ~~ sta04ec12 
  lonelye18 ~~ lonelye18 # Residual variances
  sicountrye18 ~~ sicountrye18 
  lonelysm2e ~~ lonelysm2e 
  psstdysm2er ~~ psstdysm2er 
'
```
 
```{r fit CLPM3}
CLPM3.fit <- lavaan(CLPM3, 
                   data = dat, 
                   missing = 'ML',
                   meanstructure = TRUE, 
                   int.ov.free = TRUE,
                   se = "robust",
                   estimator = "MLR" #maximum likelihood with robust (Huber-White) standard errors and a scaled (Yuan-Bentler) and robust test statistic 
                   ) 

CLPM3.fit.summary <- summary(CLPM3.fit, 
                            fit.measures = TRUE,
                            standardized = TRUE)

#Table of model fit 
CLPM3.fit.summary.fit <- table.model.fit(CLPM3.fit.summary)
CLPM3.fit.summary.fit

#Table of regression coefficients and covariances (concurrent associations)
CLPM3.fit.summary.reg <- table.model.coef(model = CLPM3.fit.summary, type = "CLPM", constraints = "Yes")
CLPM3.fit.summary.reg
```

```{r LRT for CLPMcomb and CLPMcomb2}
lavTestLRT(CLPM.fit, CLPM3.fit, method = "satorra.bentler.2010")
```
Significantly worse fit (p=0.022)- can't constrain autoregressive + crosslag paths. 

*Constrained autoregressive (CLPM2) as the chosen model.*

# RI-CLPM

```{r specify RICLPM}
RICLPM <- '
  # Create between components (random intercepts treated as factors here)
  RIln =~ 1*lonelye12 + 1*lonelye18 + 1*lonelysm2e     #x
  RIss =~ 1*sta04ec12 + 1*sicountrye18 + 1*psstdysm2er #y

  # Create within-person centered variables
  wln12 =~ 1*lonelye12
  wln18 =~ 1*lonelye18
  wln26 =~ 1*lonelysm2e 
  wss12 =~ 1*sta04ec12
  wss18 =~ 1*sicountrye18
  wss26 =~ 1*psstdysm2er
  
  # Estimate the lagged effects between the within-person centered variables
  wln18 + wss18 ~ wln12 + wss12
  wln26 + wss26 ~ wln18 + wss18 
  
  # Estimate the covariance between the within-person centered variables at the first wave
  wln12 ~~ wss12 # Covariance
  
  # Estimate the covariances between the residuals of the within-person centered variables (the innovations)
  wln18 ~~ wss18
  wln26 ~~ wss26
  
  # Estimate the variance and covariance of the random intercepts
  RIln ~~ RIln
  RIss ~~ RIss
  RIln ~~ RIss
  
  # Estimate the (residual) variance of the within-person centered variables.
  wln12 ~~ wln12 # Variances
  wss12 ~~ wss12 
  wln18 ~~ wln18 # Residual variances
  wss18 ~~ wss18 
  wln26 ~~ wln26 
  wss26 ~~ wss26 
'
```

```{r fit RICLPM}
RICLPM.fit <- lavaan(RICLPM,               # model
                     data = dat,           # data
                     missing = 'ML',       # how to handle missing data 
                     meanstructure = TRUE, # adds intercepts/means to the model for both observed and latent variables
                     se = "robust",        # robust standard errors
                     int.ov.free = TRUE,   # if FALSE, the intercepts of the observed variables are fixed to zero
                     estimator = "MLR" #maximum likelihood with robust (Huber-White) standard errors and a scaled (Yuan-Bentler) and robust test statistic
)

RICLPM.fit.summary <- summary(RICLPM.fit, 
                              fit.measures = TRUE,
                              standardized = TRUE)

#Table of model fit 
RICLPM.fit.summary.fit <- table.model.fit(RICLPM.fit.summary)
#Table of regression coefficients and covariances (concurrent associations)
RICLPM.fit.summary.reg <- table.model.coef(model = RICLPM.fit.summary, type = "RICLPM", constraints = "No")
RICLPM.fit.summary.reg
```

# Constrained RI-CLPM - autoregressive + cross-lag

```{r specify RICLPM2}
RICLPM2 <- '
  # Create between components (random intercepts treated as factors here)
  RIln =~ 1*lonelye12 + 1*lonelye18 + 1*lonelysm2e     #x
  RIss =~ 1*sta04ec12 + 1*sicountrye18 + 1*psstdysm2er #y

  # Create within-person centered variables
  wln12 =~ 1*lonelye12
  wln18 =~ 1*lonelye18
  wln26 =~ 1*lonelysm2e 
  wss12 =~ 1*sta04ec12
  wss18 =~ 1*sicountrye18
  wss26 =~ 1*psstdysm2er
  
  # Estimate the lagged effects between the within-person centered variables
  wln18 ~ a*wln12 + c*wss12
  wss18 ~ d*wln12 + b*wss12
  
  wln26 ~ a*wln18 + c*wss18 
  wss26 ~ d*wln18 + b*wss18 
  
  # Estimate the covariance between the within-person centered variables at the first wave
  wln12 ~~ wss12 # Covariance
  
  # Estimate the covariances between the residuals of the within-person centered variables (the innovations)
  wln18 ~~ wss18
  wln26 ~~ wss26
  
  # Estimate the variance and covariance of the random intercepts
  RIln ~~ RIln
  RIss ~~ RIss
  RIln ~~ RIss
  
  # Estimate the (residual) variance of the within-person centered variables.
  wln12 ~~ wln12 # Variances
  wss12 ~~ wss12 
  wln18 ~~ wln18 # Residual variances
  wss18 ~~ wss18 
  wln26 ~~ wln26 
  wss26 ~~ wss26 
'
```

```{r fit RICLPM2}
RICLPM2.fit <- lavaan(RICLPM2,               # model
                     data = dat,           # data
                     missing = 'ML',       # how to handle missing data 
                     meanstructure = TRUE, # adds intercepts/means to the model for both observed and latent variables
                     se = "robust",        # robust standard errors
                     int.ov.free = TRUE,   # if FALSE, the intercepts of the observed variables are fixed to zero
                     estimator = "MLR" #maximum likelihood with robust (Huber-White) standard errors and a scaled (Yuan-Bentler) and robust test statistic
)

RICLPM2.fit.summary <- summary(RICLPM2.fit, 
                              fit.measures = TRUE,
                              standardized = TRUE)

#Table of model fit 
RICLPM2.fit.summary.fit <- table.model.fit(RICLPM2.fit.summary)
#Table of regression coefficients and covariances (concurrent associations)
RICLPM2.fit.summary.reg <- table.model.coef(model = RICLPM2.fit.summary, type = "RICLPM", constraints = "Yes")
RICLPM2.fit.summary.reg
```

```{r LRT for RICLPM and RICLPM2}
lavTestLRT(RICLPM.fit, RICLPM2.fit, method = "satorra.bentler.2010")
```

No significant loss of fit when constrained - go with constrained model. 